Internship Summary: Mid-Term Review

### 2020-07-16 RapportStage/Rapport.tex
update wednesday afternoon
update

Within perturbation theory, the Schr\"odinger equation is usually rewritten as 
\begin{equation}
	\bH(\lambda) \Psi = (\bH^{(0)} + \lambda \bV ) \Psi(\lambda) = E(\lambda) \Psi,
\end{equation}
with
\begin{equation}
\bV=\bH - \bH^{(0)}
\end{equation}
The energy can then be written as a power series of $\lambda$
\begin{equation}
	E(\lambda) = \sum_{k=0}^\infty \lambda^k E^{(k)}
\end{equation}
where $\lambda$ is a coupling parameter set equal to 1 at the end of the calculation. However it is not guaranteed that the series $E(\lambda)$ has a radius of convergence $\abs{\lambda_0} < 1$. It means that the series is divergent for the physical system at $\lambda=1$. One can prove that $\abs{\lambda_0}$ can be obtained by extending $\lambda$ in the complex plane and looking for the singularities of $E(\lambda)$. This is due to the following theorem \cite{Goodson_2011}: The Taylor series about a point $z_0$ of a function over the complex $z$ plane will converge at a value $z_1$ if the function is non-singular at all values of $z$ in the circular region centered at $z_0$ with radius $\abs{z_1 − z_0}$. If the function has a singular point $z_s$ such that $\abs{z_s − z_0} < \abs{z_1 − z_0}$, then the series will diverge when evaluated at $z_1$. This theorem means that the radius of convergence of the perturbation series is equal to distance to the origin of the closest singularity of $E(\lambda)$.

@@

The presence of an EP close to the real axis is characteristic of a sharp avoided crossings. Yet at such an avoided crossings eigenstates change abruptly. Although it is now well understood that EPs are closely related to quantum phase transitions, the link between the type of QPT (ground state or excited state, first or superior order) and EPs still need to be clarify. One of the major challenge in order to do this reside in our ability to compute the distribution of EPs. The numerical assignment of an EP to two energies on the real axis is very difficult in large dimensions. Cejnar et al. developped a method based on a Coulomb analogy giving access to the density of EP close to the real axis \cite{Cejnar_2005, Cejnar_2007}. More recently Stransky and co-workers proved that the distribution of EPs is not the same around a QPT of first or second order \cite{Stransky_2018}. Moreover, that when the dimension of the system increases they tends towards the real axis in a different manner, meaning respectively exponentially and algebraically.

It seems like our understanding of the physics of spatial and/or spin symmetry breaking in the Hartree-Fock theory can be enlightened by quantum phase transition theory. Indeed, the second derivative of the energy is discontinuous at the Coulson-Fischer point which mean that the system undergo a second order quantum phase transition. The $\beta$ singularities introduced by Sergeev to describe the EPs modeling the formation of a bound cluster of electrons are actually a more general class of singularities.

@@

@@

@@

\section{Perturbation theory and exceptional points}

>>>>

Within perturbation theory, the Schr\"odinger equation is usually rewritten as 
\begin{equation}
	\bH(\lambda) \Psi = (\bH^{(0)} + \lambda \bV ) \Psi(\lambda) = E(\lambda) \Psi,
\end{equation}
with
\begin{equation}
\bV=\bH - \bH^{(0)}
\end{equation}
The energy can then be written as a power series of $\lambda$
\begin{equation}
	E(\lambda) = \sum_{k=0}^\infty \lambda^k E^{(k)}
\end{equation}
where $\lambda$ is a coupling parameter set equal to 1 at the end of the calculation. However it is not guaranteed that the series $E(\lambda)$ has a radius of convergence $\abs{\lambda_0} < 1$. It means that the series is divergent for the physical system at $\lambda=1$. One can prove that $\abs{\lambda_0}$ can be obtained by extending $\lambda$ in the complex plane and looking for the singularities of $E(\lambda)$. This is due to the following theorem \cite{Goodson_2012}: The Taylor series about a point $z_0$ of a function over the complex $z$ plane will converge at a value $z_1$ if the function is non-singular at all values of $z$ in the circular region centered at $z_0$ with radius $\abs{z_1 − z_0}$. If the function has a singular point $z_s$ such that $\abs{z_s − z_0} < \abs{z_1 − z_0}$, then the series will diverge when evaluated at $z_1$. This theorem means that the radius of convergence of the perturbation series is equal to distance to the origin of the closest singularity of $E(\lambda)$.

@@

The presence of an EP close to the real axis is characteristic of a sharp avoided crossing. Yet at such an avoided crossing eigen states change abruptly. Although it is now well understood that EPs are closely related to quantum phase transitions, the link between the type of QPT (ground state or excited state, first or superior order) and EPs still need to be clarify. One of the major challenge in order to do this reside in our ability to compute the distribution of EPs. The numerical assignment of an EP to two energies on the real axis is very difficult in large dimensions. Cejnar et al. developed a method based on a Coulomb analogy giving access to the density of EP close to the real axis \cite{Cejnar_2005, Cejnar_2007}. More recently Stransky and co-workers proved that the distribution of EPs is not the same around a QPT of first or second order \cite{Stransky_2018}. Moreover, that when the dimension of the system increases they tends towards the real axis in a different manner, meaning respectively exponentially and algebraically.

It seems like our understanding of the physics of spatial and/or spin symmetry breaking in the Hartree-Fock theory can be enlightened by quantum phase transition theory. Indeed, the second derivative of the energy is discontinuous at the Coulson-Fischer point which means that the system undergo a second order quantum phase transition. The $\beta$ singularities introduced by Sergeev to describe the EPs modeling the formation of a bound cluster of electrons are actually a more general class of singularities.

@@

The minimization of the UHF and RHF ground state energies with respect to $\chi$ respectively yiels the following equations:

\begin{equation}
\frac{2(75+6R-56R\cos(2\chi))\sin(2\chi))}{75R^2}=0
\label{eq:uhfstatio}
\end{equation}

\begin{equation}
\frac{2(75+6R+44R\cos(2\chi))\sin(2\chi))}{75R^2}=0
\label{eq:rhfstatio}
\end{equation}

It is evident that those equations are both verified for $\chi=\frac{n\pi}{2}$ with $n$ an integer. Those value of $\chi$ lead to the

@@

Apparition of the sb-UHF solution
Negative R possible, sb-UHF now a maximum
Less know sb-RHF solution, maximum and minimum inversed
Charge density wave vs Spin density wave

NRJ graphics

@@

\section{Radius of convergence and exceptional points}

\subsection{Evolution of the radius of convergence}

Different partitioning

$Y_{l0}$ vs $P_l(\cos(\theta))$

Size of the basis set

Strong coupling ???

\subsection{Exceptional points}

RHF vs UHF

UHF spin contamination -> Riemann surfaces ??

UHF: CF point = QPT, ESQPT ???

PT broken symmetry sb UHF
